\section{Continuous Random Variables}

\subsection{Continuous Random Variables}

\subsubsection*{Continuous Random Variables}
\begin{itemize}
    \item A random variable with a continuous distribution($X_S$)
    \item A r.v. has a continuous distribution if its CDF($F_X=P(X\leq x)$) is
    \begin{itemize}
        \item Differentiable except finitely many points
        \item Continuous everywhere
    \end{itemize}
\end{itemize}

\subsubsection*{Probability Density Function}
\begin{itemize}
    \item Probability density function(PDF) of continuous r.v. $X$ is a derivative of its CDF
    \begin{equation}
        f_X(x)=\frac{d}{dx}F_X(x)
    \end{equation}
    \item Support of $X$: $\{x\mid f_X(x)>0\}$
    \item PDF to CDF
    \begin{equation}
        F_X(x)=\int_{-\infty}^{x}f_X(t)~dt
    \end{equation}
    \item For a given $A\subset\mathbb{R}$,
    \begin{equation}
        P(X\in A)=\int_Af_X(x)~dx
    \end{equation}
    \begin{itemize}
        \item Generalization of \ref{eq:prp:discrete-pmf-subset}
    \end{itemize}
    \item Valid PDFs: Any PDF $f$ satisfies
    \begin{itemize}
        \item Nonnegative: $\forall x\in\mathbb{R},~f(x)\geq 0$
        \item Integrates to $1$: $\int_{-\infty}^{\infty}f(x)~dx=1$
    \end{itemize}
    \item Expectation of continuous r.v.
    \begin{equation}
        E(X)=\int_{-\infty}^{\infty}xf_X(x)~dx
    \end{equation}
    \begin{itemize}
        \item Not every dist. has a mean (e.g. Cauchy dist. $f_X(x)=\frac{1}{\pi(1+x^2)}$, $E(X)=\infty$)
    \end{itemize}
    \item Continuous LOTUS: $f$ is a PDF of cont. r.v. $X$ and $g:\mathbb{R}\to\mathbb{R}$, then for $Y=g(X)$,
    \begin{equation}
        E(Y)=E(g(X))=\int_{-\infty}^{\infty}g(x)f(x)~dx
    \end{equation}
\end{itemize}

\subsubsection*{Uniform Distribution}
\begin{equation}
    X\sim\text{Unif}(a,b)\implies f_X(x)=\begin{cases}
        \frac{1}{b-a} & x\in(a,b)\\
        0 & \text{else}
    \end{cases}
\end{equation}
\begin{itemize}
    \item Mean: $E(U)=\frac{a+b}{2}$
    \item Variance: $\text{Var}(U)=\frac{1}{12}(b-a)^2$
\end{itemize}

\subsubsection*{Universality of Uniform}
\begin{itemize}
    \item Condition: $F:\mathbb{R}\to(0,1)$ is a valid CDF and $\exists~F^{-1}:(0,1)\to\mathbb{R}$
    \begin{itemize}
        \item $\impliedby$ $F$ is continuous, strictly increasing on its support
    \end{itemize}
    \item For such $F$,
    \begin{enumerate}
        \item $U\sim\text{Unif}(0,1)$, $X=F^{-1}(U)$ $\implies$ $X$ is a r.v. with CDF $F$
        \item $X$: r.v. with CDF $F$ $\implies$ $F(X)\sim\text{Unif}(0,1)$
    \end{enumerate}
    \item We can generate a target distribution from a uniformly distributed data
    \begin{itemize}
        \item Limitation: CDF of target dist. must have its inverse
    \end{itemize}
\end{itemize}

\subsubsection*{Normal (Gaussian) Distribution}
\begin{itemize}
    \item Standard Normal Distribution
    \begin{equation}
        Z\sim\mathcal{N}(0,1)\implies \varphi(z)=\frac{1}{\sqrt{2\pi}}e^{-\frac{z^2}{2}}~~(z\in(-\infty,\infty))
    \end{equation}
    \begin{itemize}
        \item $\varphi$ is a PDF of $Z$
        \item $E(Z)=0$, $\text{Var}(Z)=1$
        \item Standard normal CDF: $\Phi(z)=\int_{-\infty}^z\varphi(t)~dt$
    \end{itemize}
    \item Normal Distribution
    \begin{itemize}
        \item $Z\sim\mathcal{N}(0,1)$ $\implies$ $X=\mu+\sigma Z\sim\mathcal{N}(\mu,\sigma^2)$
        \item $E(X)=\mu$, $\text{Var}(X)=\sigma^2$
    \end{itemize}
\end{itemize}

\subsubsection*{Exponential Distribution}
\begin{equation}
    X\sim\text{Expo}(\lambda)\implies f_X(x)=\begin{cases}
        \lambda e^{-\lambda x} & x>0 \\
        0 & x\leq 0
    \end{cases}~~(\lambda>0)
\end{equation}
\begin{itemize}
    \item CDF: $F_X(x)=1-e^{-\lambda x}$ for $x>0$
    \item $Y=\lambda_0X\sim\text{Expo}\left(\frac{\lambda}{\lambda_0}\right)$
    \item Expectation and Variance
    \begin{equation}
        E(X)=\frac{1}{\lambda},~~\text{Var}(X)=\frac{1}{\lambda^2}
    \end{equation}
    \item Exponential dist. is a \textit{continuous} version of Geometric dist.
    \begin{itemize}
        \item Parameter $\lambda>0$: Interpreted as a \textit{averaged number} of events occuring during unit time
        \item $X$: waiting time to see an event occurrence
    \end{itemize}
    \item Memoryless Property: Dist. of r.v. $X$ has \textit{memoryless property} if
    \begin{equation}
        \forall~s,t>0,~P(X\geq s+t\mid X\geq s)=P(X\geq t)
    \end{equation}
    \begin{itemize}
        \item Exponential dist. and geometric dist. do have memoryless property
    \end{itemize}
\end{itemize}
